Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-10-25 WSSSPE3/report/WSSSPE3_review.tex
edits/comment in keynote

Turk reminded the audience that the volunteer model means unpaid labor. On this
note, he recommended Ashe Dryden's blog post on the
\href{http://www.ashedryden.com/blog/the-ethics-of-unpaid-labor-and-the-oss-community}{ethics
of unpaid labor and the open source software community}. Often times, a person
funded to work full time on some scientific project can spend a small amount of
time for working on a piece of software necessary for the project. However, all
researchers' ability to participate in that volunteer community are not always
the same and may not always be aligned with the research projects. From Turk's
experience, we cannot always rely on unpaid labor and volunteer time to sustain
a piece of software---this came down to the notions of the top-down and the
bottom-up approaches, i.e., the funded versus the grassroots. However, Turk
pointed out that bottom-up volunteers driven projects can be just as large scale
as a top-down model software development.

When Turk talked about sustaining scientific software, it really meant to him as
conducting scientific inquiries by often some specific software and sustaining
the people we care about, our careers, and the future of our fields. According
to Turk, we all have an invested stake in sustaining scientific software. Again
it came down to the top-down versus bottom-up notion that we can have anointed
specific software packages to be sustained at the expense of new grassroot
developments.

To close, Turk talked about possibilities of why we want to sustain scientific
software: devotion to science and interests in pursuing the next stage of
research; fun and creative thrill in writing code and paper; usefulness with
measurable impacts, for example, LINPACK and HDF groups providing of data
storage to satellites, which goes beyond usefulness to necessity. Lastly, Turk
presented his wishlist of questions to be answered: 
%
\begin{itemize}

@@

\item how do we know when it is time to end some software and move on? For
example, shall we stop sustaining Python and switch to Julia and Javascript?

>>>>

Turk reminded the audience that the volunteer model means unpaid labor. On this
note, he recommended Ashe Dryden's blog post on the
ethics of unpaid labor and the open source software
community\footnote{http://www.ashedryden.com/blog/the-ethics-of-unpaid-labor-and-the-oss-community}.
Often times, a person
funded to work full time on a scientific project can spend a small amount of
time for working on a piece of software necessary for that project. However, all
researchers' ability to participate in that volunteer community are not always
the same and may not always be aligned with their research projects. From Turk's
experience, we cannot always rely on unpaid labor and volunteer time to sustain
a piece of software---this came down to the notions of the top-down and the
bottom-up approaches, i.e., the funded versus the grassroots. However, Turk
pointed out that bottom-up, volunteer-driven projects can be just as large
as a top-down software development project.

Turk said that sustaining scientific software really meant to him
conducting scientific inquiries, often by some specific software, and sustaining
the people we care about, our careers, and the future of our fields. According
to Turk, we all have an invested stake in sustaining scientific software. Again
it came down to the top-down versus bottom-up notion that we can have anointed
specific software packages to be sustained at the expense of new grassroots
developments. \katznote{I don't understand this last sentence.}

Turk talked about possible reasons why we want to sustain scientific
software: devotion to science and interests in pursuing the next stage of
research; fun and creative thrill in writing codes and papers; usefulness with
measurable impacts, for example, LINPACK and HDF groups providing of data
storage to satellites, which goes beyond usefulness to necessity. Lastly, Turk
presented his wishlist of questions to be answered: 
%
\begin{itemize}

@@

\item How do we know when it is time to end some software and move on? For
example, should we stop sustaining Python and switch to Julia and Javascript?
